---
title: "ClimateNA"
author: "Jen Baron, j.baron@alumni.ubc.ca, UBC Tree Ring Lab"
date: 'June 17, 2020'
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
---

# Packages

```{r}
library(ggplot2)
library(dplyr)
library(purrr)
library(tidyr)
library(gridExtra)
```

# Load Data

```{r}
monthly <- read.csv("data/output_1901-2018M.csv")
season <- read.csv("data/output_1901-2018S.csv")
annual <- read.csv("data/output_1901-2018Y.csv")
all <- read.csv("data/output_1901-2018MSY.csv")
```

# Prepare Data





## Annual

Directly calculated annual variables:

- MAT	 	mean annual temperature (°C),
- MWMT 	mean warmest month temperature (°C),
- MCMT  	mean coldest month temperature (°C),
- TD 		temperature difference between MWMT and MCMT, or continentality (°C),
- MAP 		mean annual precipitation (mm),
- MSP 		May to September precipitation (mm),
- AHM  	annual heat-moisture index (MAT+10)/(MAP/1000))
- SHM 	 	summer heat-moisture index ((MWMT)/(MSP/1000))
	
Derived annual variables:

- DD<0		degree-days below 0°C, chilling degree-days
- DD>5		degree-days above 5°C, growing degree-days
- DD<18	degree-days below 18°C, heating degree-days
- DD>18	degree-days above 18°C, cooling degree-days
- NFFD		the number of frost-free days
- FFP		frost-free period
- bFFP		the day of the year on which FFP begins
- eFFP		the day of the year on which FFP ends
- PAS	precipitation as snow (mm). For individual years, it covers the period between August in the previous year and July in the current year.
- EMT		extreme minimum temperature over 30 years
- EXT		extreme maximum temperature over 30 years
- Eref		Hargreaves reference evaporation (mm)
- CMD		Hargreaves climatic moisture deficit (mm)
- MAR		mean annual solar radiation (MJ m‐2 d‐1)
- RH		mean annual relative humidity (%)

```{r}
str(annual)
```



## Seasonal

Seasonal variables:

- Winter (_wt): Dec (prev. yr for an individual year) - Feb for annual, Jan, Feb, Dec for normals
- Spring (_sp): March, April and May
- Summer (_sm): June, July and August
- Autumn (_at): September, October and November 

Directly calculated seasonal variables:

-  Tave_wt	winter mean temperature (°C)
-  Tmax_wt	winter mean maximum temperature (°C)
-  Tmin_wt	winter mean minimum temperature (°C)
-  PPT_wt	winter precipitation (mm)
-  RAD_wt	winter solar radiation (MJ m‐2 d‐1)

Derived seasonal variables:
- DD_0_wt	winter degree-days below 0°C
- DD5_wt	winter degree-days below 5°C
- DD_18_wt	winter degree-days below 18°C
- DD18_wt	winter degree-days below 18°C
- NFFD_wt 	winter number of frost-free days
- PAS_wt 	winter precipitation as snow (mm)
- Eref_wt 	winter Hargreaves reference evaporation (mm)
- CMD_wt 	winter Hargreaves climatic moisture deficit (mm)
- RH_wt		winter relative humidity (%)


```{r}
str(season)
```

```{r}
season_Tmax <- season %>% 
  select(Year, starts_with("Tmax")) %>%
  gather(key = "Season", value = "Tmax", 2:5)
season_Tmax$Season <- as.factor(season_Tmax$Season)
levels(season_Tmax$Season) <- c("Autumn","Summer","Spring","Winter")

season_Tmin <- season %>% 
  select(Year, starts_with("Tmin")) %>%
  gather(key = "Season", value = "Tmin", 2:5)
season_Tmin$Season <- as.factor(season_Tmin$Season)
levels(season_Tmin$Season) <- c("Autumn","Summer","Spring","Winter")

season_Tave <- season %>% 
  select(Year, starts_with("Tave")) %>%
  gather(key = "Season", value = "Tave", 2:5)
season_Tave$Season <- as.factor(season_Tave$Season)
levels(season_Tave$Season) <- c("Autumn","Summer","Spring","Winter")

season_PPT <- season %>% 
  select(Year, starts_with("PPT")) %>%
  gather(key = "Season", value = "PPT", 2:5)
season_PPT$Season <- as.factor(season_PPT$Season)
levels(season_PPT$Season) <- c("Autumn","Summer","Spring","Winter")

season_Rad <- season %>% 
  select(Year, starts_with("Rad")) %>%
  gather(key = "Season", value = "Rad", 2:5)
season_Rad$Season <- as.factor(season_Rad$Season)
levels(season_Rad$Season) <- c("Autumn","Summer","Spring","Winter")

season_Eref <- season %>% 
  select(Year, starts_with("Eref")) %>%
  gather(key = "Season", value = "Eref", 2:5)
season_Eref$Season <- as.factor(season_Eref$Season)
levels(season_Eref$Season) <- c("Autumn","Summer","Spring","Winter")

season_CMD <- season %>% 
  select(Year, starts_with("CMD")) %>%
  gather(key = "Season", value = "CMD", 2:5)
season_CMD$Season <- as.factor(season_CMD$Season)
levels(season_CMD$Season) <- c("Autumn","Summer","Spring","Winter")

season_RH <- season %>% 
  select(Year, starts_with("RH")) %>%
  gather(key = "Season", value = "RH", 2:5)
season_RH$Season <- as.factor(season_RH$Season)
levels(season_RH$Season) <- c("Autumn","Summer","Spring","Winter")
```

```{r}
season_tidy <-list(season_Tmax, season_Tmin, season_Tave, season_PPT, season_Rad, season_Eref, season_CMD, season_RH) %>% reduce(full_join)

str(season_tidy)
```



## Monthly

Primary monthly variables:

- Tave01 – Tave12 	January - December mean temperatures (°C)
- TMX01 – TMX12 	January - December maximum mean temperatures (°C)
- TMN01 – TMN12 	January - December minimum mean temperatures (°C)
- PPT01 – PPT12  	January - December precipitation (mm)
- RAD01 – RAD12  	January - December solar radiation (MJ m‐2 d‐1)

Derived monthly variables:
- DD_0_01 – DD_0_12   January - December degree-days below 0°C
- DD5_01 – DD5_12     January - December degree-days above 5°C
- DD_18_01 – DD_18_12  	January - December degree-days below 18°C
- DD18_01 – DD18_12    	January - December degree-days above 18°C
- NFFD01 – NFFD12 	    	January - December number of frost-free days
- PAS01 – PAS12 		January – December precipitation as snow (mm)
- Eref01 – Eref12 	January – December Hargreaves reference evaporation (mm)
- CMD01 – CMD12 	January – December Hargreaves climatic moisture deficit (mm)
- RH01 – RH12		January – December relative humidity (%)


```{r}
str(monthly)
```


```{r}
monthly_Tmax <- monthly %>% 
  select(Year, starts_with("Tmax")) %>%
  gather(key = "Month", value = "Tmax", 2:13)
monthly_Tmax$Month <- as.factor(monthly_Tmax$Month)
levels(monthly_Tmax$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_Tmin <- monthly %>% 
  select(Year, starts_with("Tmin")) %>%
  gather(key = "Month", value = "Tmin", 2:13)
monthly_Tmin$Month <- as.factor(monthly_Tmin$Month)
levels(monthly_Tmin$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_Tave <- monthly %>% 
  select(Year, starts_with("Tave")) %>%
  gather(key = "Month", value = "Tave", 2:13)
monthly_Tave$Month <- as.factor(monthly_Tave$Month)
levels(monthly_Tave$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_PPT <- monthly %>% 
  select(Year, starts_with("PPT")) %>%
  gather(key = "Month", value = "PPT", 2:13)
monthly_PPT$Month <- as.factor(monthly_PPT$Month)
levels(monthly_PPT$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_Rad <- monthly %>% 
  select(Year, starts_with("Rad")) %>%
  gather(key = "Month", value = "Rad", 2:13)
monthly_Rad$Month <- as.factor(monthly_Rad$Month)
levels(monthly_Rad$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_Eref <- monthly %>% 
  select(Year, starts_with("Eref")) %>%
  gather(key = "Month", value = "Eref", 2:13)
monthly_Eref$Month <- as.factor(monthly_Eref$Month)
levels(monthly_Eref$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_CMD <- monthly %>% 
  select(Year, starts_with("CMD")) %>%
  gather(key = "Month", value = "CMD", 2:13)
monthly_CMD$Month <- as.factor(monthly_CMD$Month)
levels(monthly_CMD$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

monthly_RH <- monthly %>% 
  select(Year, starts_with("RH")) %>%
  gather(key = "Month", value = "RH", 2:13)
monthly_RH$Month <- as.factor(monthly_RH$Month)
levels(monthly_RH$Month) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
```


```{r}
monthly_tidy <- list(monthly_Tmax, monthly_Tmin, monthly_Tave, monthly_PPT, monthly_Rad, monthly_Eref, monthly_CMD, monthly_RH) %>% reduce(full_join)

str(monthly_tidy)
```


# Visualize Data



## Annual

```{r}
 ggplot() +
  geom_line(data=annual, aes(x=Year, y=MAT), colour = "darkgreen") +
  xlab("Year") +
  ylab("Mean Annual Temperature (°C)") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot() +
  geom_line(data=annual, aes(x=Year, y=MAP), colour = "darkgreen") +
  #Aesthetics
  xlab("Year") +
  ylab("Mean Annual Precipitation (mm)") +
  scale_x_continuous(breaks=seq(1900,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot() +
  geom_line(data=annual, aes(x=Year, y=MSP), colour = "darkgreen") +
  xlab("Year") +
  ylab("May to September Precipitation (mm)") +
  scale_x_continuous(breaks=seq(1900,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot() +
  geom_line(data=annual, aes(x=Year, y=PAS), colour = "darkgreen") +
  xlab("Year") +
  ylab("Precipitation as Snow (mm)") +
  scale_x_continuous(breaks=seq(1900,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot() +
  geom_line(data=annual, aes(x=Year, y=CMD), colour = "darkgreen") +
  xlab("Year") +
  ylab("Climate Moisture Defecit") +
  scale_x_continuous(breaks=seq(1900,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```

## Seasonal
```{r}
str(season_tidy)
```


```{r}
ggplot (data = season_tidy, aes(x=Year, y = Tave, col= Season)) +
  geom_line() +
  theme_classic()

ggplot (data = season_tidy, aes(x=Year, y = PPT, col= Season)) +
  geom_line() +
  facet_grid(Season~.) +
  theme_classic()

ggplot (data = season_tidy, aes(x=Year, y = CMD, col= Season)) +
  geom_line() +
  theme_classic()
```


```{r}
str(season_tidy)
```


```{r}
w_tempNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Winter")), aes(x = Year, y = Tave)) +
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_tempNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Spring")), aes(x = Year, y = Tave)) +
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")

su_tempNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Summer")), aes(x = Year, y = Tave)) +
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

au_tempNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Autumn")), aes(x = Year, y = Tave)) +
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Autumn") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

seasonal_temp_plotNA <- grid.arrange(w_tempNA, sp_tempNA, su_tempNA, au_tempNA, nrow = 2, ncol = 2)
```


```{r}
w_precipNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Winter")), aes(x = Year, y = PPT)) +
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_precipNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Spring")), aes(x = Year, y = PPT)) +
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")

su_precipNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Summer")), aes(x = Year, y = PPT)) +
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

au_precipNA <- ggplot() +
  geom_line(data=(season_tidy %>% filter(Season == "Autumn")), aes(x = Year, y = PPT)) +
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Autumn") +
  scale_x_continuous(limits = c(1900, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

seasonal_precip_plotNA <- grid.arrange(w_precipNA, sp_precipNA, su_precipNA, au_precipNA, nrow = 2, ncol = 2)
```



## Monthly

```{r}
str(monthly_tidy)
```




```{r}
ggplot() + 
  geom_boxplot(data = monthly_tidy, aes(x = Month, y = Tave), alpha = 0.3) +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot() + 
  geom_boxplot(data = monthly_tidy, aes(x = Month, y = PPT), alpha = 0.7) + 
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 


ggplot() + 
  geom_boxplot(data = monthly_tidy, aes(x = Month, y = CMD), alpha = 0.7) + 
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 


ggplot() + 
  geom_boxplot(data = monthly_tidy, aes(x = Month, y = RH), alpha = 0.7) + 
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

```




```{r}
ggplot(data = monthly_tidy, aes(x=Year, y = Tave)) +
  geom_line() +
  facet_grid(Month~.) + 
  theme_classic() 

ggplot(data = monthly_tidy, aes(x=Year, y = PPT)) +
  geom_line() +
  facet_grid(Month~.) + 
  theme_classic() 

ggplot(data = monthly_tidy, aes(x=Year, y = CMD)) +
  geom_line() +
  facet_grid(Month~.) + 
  theme_classic() 
```

### Climograph

```{r}
NA_p <- monthly_tidy %>% 
  group_by(Month) %>%
  summarize(mean_precip= mean(PPT), sd_precip = sd(PPT), n = n()) #%>%
  
NA_p$Month <- as.numeric(NA_p$Month)

NA_p <- NA_p %>% mutate(se_precip = sd_precip / sqrt(n), lower = mean_precip - se_precip, upper = mean_precip + se_precip) %>% round(2)
NA_p
```

```{r}
ggplot() +
  geom_boxplot(data = monthly_tidy, aes(x = Month, y = Tave), fill = "red3", alpha = 0.7) +
  geom_line(data = NA_p, aes(x = Month, y = mean_precip-75), col= "steelblue4", size = 1) +
    geom_ribbon(data = NA_p, aes(x = Month, ymin = (lower-75), ymax = (upper-75)), alpha = 0.3, fill= "steelblue4") +
  scale_y_continuous("Temperature (°C)", sec.axis = sec_axis(~ . + 75, name = "Precipitation (mm)")) +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```





# Save Outputs

```{r}
write.csv(monthly_tidy, "outputs/monthly_tidy.csv")
write.csv(season_tidy, "outputs/season_tidy.csv")
```


# Reproducibility
```{r}
Sys.time()
git2r::repository()
sessionInfo()
```