---
title: "Instrumental Climate Data"
author: "Jen Baron, j.baron@alumni.ubc.ca, UBC Tree Ring Lab"
date: 'August 8, 2019'
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
---


# Packages
```{r}
library(ggplot2) #for plotting
library(tidyverse)
library(dplyr)
library(GGally)
library(gridExtra)
```


# Read in Data

The Pinery dataset was compiled in excel from monthly mean temperature and total precipitation data using open access Environment Canada records for stations within 30km of the Pinery with at least 10 consecutive years of data (http://climate.weather.gc.ca/historical_data/search_historic_data_e.html). 

The London datasets are homogenized data from Environment Canada that were converted to tidy format in excel. 

```{r}
pinery <- read.csv("data/pinery_climate.csv")
pinery$station <- as.factor(pinery$station) #Make a factor 
str(pinery)
london_precip <- read.csv("data/london_precip.csv")
london_temp <- read.csv("data/london_temp.csv")
london_precip$month <- as.factor(london_precip$month) #Make a factor 
london_temp$month <- as.factor(london_temp$month) #Make a factor 
```

```{r}
climateNA <- read_csv("data/climateNA_monthly.csv") 
climateNA.temp <- climateNA %>% select(Year, Month, Tave)
climateNA.precip <- climateNA %>% select(Year, Month, PPT)
```


# Create Regional Data

## Clean Data

Here I go through each station for the temp and precip data, removing years where all 12 months aren't present so that I can later calculate the mean values for each year. I did this manually a while ago and there is probably a more efficient way to do it, however it works and it's already done so I left it as is. 

```{r}
#Arkona
Arkona <- pinery %>% filter(station == "Arkona") #Filter out station of interest

Arkona_T <- Arkona %>% select(-total_precip) %>% filter(!is.na(Arkona$mean_temp)==T) #remove precipitation data and "NA"
table(Arkona_T$year) #Generate table after cleaning
Arkona_T2 <- Arkona_T %>% filter(year != 1882 & year != 1883 & 
                                   year != 1884 & year != 1893 & year != 1900 & year != 1902 & year != 1915) #remove years with < 12 months
table(Arkona_T2$year) #Generate table after cleaning

Arkona_P <- Arkona %>% select(-mean_temp) %>% filter(!is.na(Arkona$total_precip)==T) #remove temperature data and "NA"
#table(Arkona_P$year) #show how many observations in each year
Arkona_P2 <- Arkona_P %>% filter(year != 1882 & year != 1893 & year != 1902 & year != 1915)
#table(Arkona_P2$year) #Generate table after cleaning

#Centralia A
CentraliaA <- pinery %>% filter(station == "Centralia A")

CentraliaA_T <- CentraliaA %>% select(-total_precip) %>% filter(!is.na(CentraliaA$mean_temp)==T) #remove "NA"
#table(CentraliaA_T$year)  
CentraliaA_T2 <- CentraliaA_T %>% filter(year != 1945 & year != 1946 & year != 1947 & year != 1966)
#table(CentraliaA_T2$year)  

CentraliaA_P <- CentraliaA %>% select(-mean_temp) %>% filter(!is.na(CentraliaA$total_precip)==T) #remove "NA"
#table(CentraliaA_P$year)  
CentraliaA_P2 <- CentraliaA_P %>% filter(year != 1946 & year != 1947 & year != 1966)
#table(CentraliaA_P2$year)  

#Exeter
Exeter <- pinery %>% filter(station == "Exeter")

Exeter_T <- Exeter %>% select(-total_precip) %>% filter(!is.na(Exeter$mean_temp)==T) #remove "NA"
#table(Exeter_T$year)  
Exeter_T2 <- Exeter_T %>% filter(year != 1967 & year != 1976 & year != 1985)
#table(Exeter_T2$year)  

Exeter_P <- Exeter %>% select(-mean_temp) %>% filter(!is.na(Exeter$total_precip)==T) #remove "NA"
#table(Exeter_P$year)  
Exeter_P2 <- Exeter_P %>% filter(year != 1961 & year != 1962 & year != 1964 & year != 1976 & year != 1985)
#table(Exeter_P2$year) 

#Forest
Forest <- pinery %>% filter(station == "Forest")

Forest_T <- Forest %>% select(-total_precip) %>% filter(!is.na(Forest$mean_temp)==T) #remove "NA"
#table(Forest_T$year)  
Forest_T2 <- Forest_T %>% filter(year != 1924 & year != 1940 & year != 1962 & year != 1963)
#table(Forest_T2$year) 

Forest_P <- Forest %>% select(-mean_temp) %>% filter(!is.na(Forest$total_precip)==T) #remove "NA"
#table(Forest_P$year)  
Forest_P2 <- Forest_P %>% filter(year > 1933 & year != 1937 & year!= 1943 & year != 1955 & year != 1960 & year < 1960)
#table(Forest_P2$year) 

#Lucan
Lucan <- pinery %>% filter(station == "Lucan")

Lucan_T <- Lucan %>% select(-total_precip) %>% filter(!is.na(Lucan$mean_temp)==T) #remove "NA"
#table(Lucan_T$year)  
Lucan_T2 <- Lucan_T %>% filter(year != 1915 & year != 1924 & year != 1948 & year != 1949 & year != 1950
                               & year != 1951 & year != 1952 & year != 1953 & year != 1954 & year != 1955
                               & year != 1956 & year != 1957 & year != 1958 & year != 1959 & year != 1961)
#table(Lucan_T2$year) 

Lucan_P <- Lucan %>% select(-mean_temp) %>% filter(!is.na(Lucan$total_precip)==T) #remove "NA"
#table(Lucan_P$year)  
Lucan_P2 <- Lucan_P %>% filter(year != 1915 & year != 1923 & year != 1943 & year < 1948)
#table(Lucan_P2$year)

#Pinery
Pinery <- pinery %>% filter(station == "Pinery")

Pinery_T <- Pinery %>% select(-total_precip) %>% filter(!is.na(Pinery$mean_temp)==T) #remove "NA"
#table(Pinery_T$year)  
Pinery_T2 <- Pinery_T %>% filter(year != 1979 & year != 1984)
#table(Pinery_T2$year) 

Pinery_P <- Pinery %>% select(-mean_temp) %>% filter(!is.na(Pinery$total_precip)==T) #remove "NA"
#table(Pinery_P$year)  
Pinery_P2 <- Pinery_P %>% filter(year != 1979 & year != 1984)
#table(Pinery_P2$year)

#Strathrow-Mullifary
Strathroy <- pinery %>% filter(station == "Strathroy-Mullifary")

Strathroy_T <- Strathroy %>% select(-total_precip) %>% filter(!is.na(Strathroy$mean_temp)==T) #remove "NA"
#table(Strathroy_T$year)  
Strathroy_T2 <- Strathroy_T %>% filter(year != 2009 & year != 2010 & year != 2013 & year != 2014 & year != 2015 & year < 2017)
#table(Strathroy_T2$year) 

Strathroy_P <- Strathroy %>% select(-mean_temp) %>% filter(!is.na(Strathroy$total_precip)==T) #remove "NA"
#table(Strathroy_P$year)  
Strathroy_P2 <- Strathroy_P %>% filter(year != 2009 & year != 2010 & year != 2012 & year != 2013 
                                       & year != 2014 & year != 2015 & year < 2017)
#table(Strathroy_P2$year)

#Thedford (no temp)
Thedford <- pinery %>% filter(station == "Thedford")

Thedford_P <- Thedford %>% select(-mean_temp) %>% filter(!is.na(Thedford$total_precip)==T) #remove "NA"
#table(Thedford_P$year)  
Thedford_P2 <- Thedford_P %>% filter(year != 1883 & year != 1897)
#table(Thedford_P2$year)

#Watford (can't use precipitation data)
Watford <- pinery %>% filter(station == "Watford")

Watford_T <- Watford %>% select(-total_precip) %>% filter(!is.na(Watford$mean_temp)==T) #remove "NA"
#table(Watford_T$year)  
Watford_T2 <- Watford_T %>% filter(year == 1927 | year == 1928)
#table(Watford_T2$year) 

Watford_P <- Watford %>% select(-mean_temp) %>% filter(!is.na(Watford$total_precip)==T) #remove "NA"
#table(Watford_P$year)  
```


Re-combine Data
```{r}
pinery_temp <- rbind(Arkona_T2, CentraliaA_T2, Exeter_T2, Forest_T2, Lucan_T2, Pinery_T2, Strathroy_T2, Watford_T2)
pinery_precip <- rbind(Arkona_P2, CentraliaA_P2, Exeter_P2, Forest_P2, Lucan_P2, Pinery_P2, Strathroy_P2, Thedford_P2)
```


Ensure the cleaning worked by plotting histograms of the number of observations (months) in each year (should only be multiples of 12)
```{r}
#histogram (temperature)
graph <- ggplot(pinery_temp, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
  geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
    geom_hline(yintercept = 24, linetype="solid", color = "red") + 
    geom_hline(yintercept = 36, linetype="solid", color = "red") + 
  theme_bw()
graph

#histogram (precipitation)
graph <- ggplot(pinery_precip, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
   geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
    geom_hline(yintercept = 24, linetype="solid", color = "red") + 
  theme_bw()
graph
```

## Create Composite Record

Calculate one mean temperature and total precipitation for each year/month where more than one station of data is available

```{r}
#Temperature
str(pinery_temp)

pinery_temp_mean <- pinery_temp %>%
  group_by(year,month) %>%
  summarise(mean_temp_year = mean(mean_temp), sd_temp = sd(mean_temp), n = n()) %>%
  mutate(se_temp = sd_temp / sqrt(n), lower = (mean_temp_year - se_temp), upper = (mean_temp_year + se_temp)) %>% round(2)
pinery_temp_mean$month <- as.factor(pinery_temp_mean$month)

#Precipitation
pinery_precip_mean <- pinery_precip %>%
  group_by(year,month) %>%
  summarise(total_precip_year = mean(total_precip), sd_precip = sd(total_precip), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip_year - sd_precip, upper = total_precip_year + sd_precip) %>% round(2)

head(pinery_precip_mean)
```

```{r}
ggplot(pinery_precip_mean, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey")

ggplot(pinery_temp_mean, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey")
```



## Fill Empty Years



### Temperature

Fill empty months-years with data from London

```{r}
p.t <- pinery_temp_mean %>% select(year, month, mean_temp_year)

l.t <- london_temp %>% filter(month == "Jan" |
                                month == "Feb" |
                                month == "Mar" |
                                month == "Apr" |
                                month == "May" |
                                month == "Jun" |
                                month == "Jul" |
                                month == "Aug" |
                                month == "Sep" |
                                month == "Oct" |
                                month == "Nov" |
                                month == "Dec") %>% 
  mutate(month = recode(month, "Jan"="1", 
                                     "Feb"="2", 
                                     "Mar"="3",
                                     "Apr"="4",
                                     "May"="5", 
                                     "Jun"="6", 
                                     "Jul"="7",
                                     "Aug"="8",
                                     "Sep"="9",
                                     "Oct"="10",
                                     "Nov"="11",
                                     "Dec"="12"))

```


```{r}
PL.temp <- full_join(l.t, p.t) %>%
  rename("pinery" = "mean_temp_year") %>%
  rename("london" = "mean_temp")

head(PL.temp)
```

Select all rows with regional data - source = Pinery
Select all rows with NA - source = London

```{r}
PL.1 <- PL.temp %>%  #use london value where pinery is absent
  filter(is.na(pinery)) %>%
  mutate(temp = london)
PL.1$source <- "London"

PL.2 <- PL.temp %>% # use pinery value where pinery is present
  filter(!is.na(pinery)) %>%
  mutate(temp = pinery)
PL.2$source <- "Pinery"

PL.temp2 <- rbind(PL.1, PL.2) %>%
  arrange(year)

(nrow(PL.2) + nrow(PL.1)) == nrow(PL.temp) #check equal observations

PL.temp2 %>% filter(is.na(temp))
```

Fill leftover NAs with climateNA values

```{r}
head(climateNA.temp)
climateNA.temp <- climateNA.temp %>%
  rename("month" = "Month") %>%
  rename("year" = "Year") %>%
  rename("climateNA" = "Tave")

climateNA.temp$month <- as.factor(climateNA.temp$month)

PL.temp3 <- full_join(PL.temp2, climateNA.temp)

PL.temp3 <- PL.temp3[, c(1:4,7,5,6)] #re-order columns

head(PL.temp3)
```


```{r}
PL.3 <- PL.temp3 %>%  
  filter(is.na(temp)) %>%
  mutate(temp = climateNA)
PL.3$source <- "ClimateNA"

PL.4 <- PL.temp3 %>% 
  filter(!is.na(temp))

PL.temp4 <- rbind(PL.3, PL.4) %>%
  arrange(year) %>% 
  filter(year > 1890) #missing values from 1890s remain

(nrow(PL.3) + nrow(PL.4)) == nrow(PL.temp3) #check equal observations

PL.temp4 %>% filter(is.na(temp)) #values still missing
```


```{r}
ggplot(PL.temp4, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
  geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
  theme_bw()
```


### Precipitation

```{r}
p.p <- pinery_precip_mean %>% select(year, month, total_precip_year)

l.p <- london_precip %>% filter(month == "Jan" |
                                month == "Feb" |
                                month == "Mar" |
                                month == "Apr" |
                                month == "May" |
                                month == "Jun" |
                                month == "Jul" |
                                month == "Aug" |
                                month == "Sep" |
                                month == "Oct" |
                                month == "Nov" |
                                month == "Dec") %>% 
  mutate(month = recode(month, "Jan"="1", 
                                     "Feb"="2", 
                                     "Mar"="3",
                                     "Apr"="4",
                                     "May"="5", 
                                     "Jun"="6", 
                                     "Jul"="7",
                                     "Aug"="8",
                                     "Sep"="9",
                                     "Oct"="10",
                                     "Nov"="11",
                                     "Dec"="12"))

```


```{r}
p.p$month <- as.factor(p.p$month)
PL.precip <- full_join(l.p, p.p) %>%
  rename("pinery" = "total_precip_year") %>%
  rename("london" = "total_precip")

```


```{r}
PL.5 <- PL.precip %>%  #use london value where pinery is absent
  filter(is.na(pinery)) %>%
  mutate(precip = london)
PL.5$source <- "London"

PL.6 <- PL.precip %>% # use pinery value where pinery is present
  filter(!is.na(pinery)) %>% 
  mutate(precip = pinery)
PL.6$source <- "Pinery"

PL.precip2 <- rbind(PL.5, PL.6) %>%
  arrange(year)

(nrow(PL.5) + nrow(PL.6)) == nrow(PL.precip2) #check equal observations

PL.precip2 %>% filter(is.na(precip)) #values still missing
```

Fill leftover NAs with climateNA values

```{r}
head(climateNA.temp)
climateNA.precip <- climateNA.precip %>%
  rename("month" = "Month") %>%
  rename("year" = "Year") %>%
  rename("climateNA" = "PPT")

climateNA.precip$month <- as.factor(climateNA.precip$month)

PL.precip3 <- full_join(PL.precip2, climateNA.precip)

PL.precip3 <- PL.precip3[, c(1:4,7,5,6)] #re-order columns

head(PL.temp3)
```


```{r}
PL.7 <- PL.precip3 %>%  
  filter(is.na(precip)) %>%
  mutate(precip = climateNA)
PL.7$source <- "ClimateNA"

PL.8 <- PL.precip3 %>% 
  filter(!is.na(precip))

PL.7
PL.precip4 <- rbind(PL.7, PL.8) %>%
  arrange(year) %>%
  filter(year > 1890)

(nrow(PL.7) + nrow(PL.8)) == nrow(PL.precip3) #check equal observations

PL.precip4 %>% filter(is.na(precip)) #values still missing
```


```{r}
ggplot(PL.precip4, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
  geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
  theme_bw()
```

### Combine

Fix levels
```{r}
PL.temp4 <- droplevels(PL.temp4) #drops un-used levels
PL.temp4$month <- factor(PL.temp4$month, levels = c("1","2","3","4", "5", "6", "7", "8", "9", "10", "11", "12")) #re-level
levels(PL.temp4$month)

PL.precip4 <- droplevels(PL.precip4) #drops un-used levels
PL.precip4$month <- factor(PL.precip4$month, levels = c("1","2","3","4", "5", "6", "7", "8", "9", "10", "11", "12")) #re-level
levels(PL.precip4$month)
```


```{r}
PL.temp.only <- PL.temp4 %>% select(year, month, temp)
PL.precip.only <- PL.precip4 %>% select(year, month, precip)

PL.clean <- full_join(PL.temp.only, PL.precip.only, by = c("year", "month"))

str(PL.clean)

PL.clean %>% filter(is.na(temp)) #values still missing

nrow(PL.clean) == (2019-1891)*12 #check 12 months / year for the 128 years
```



# Monthly Data

## Climograph


```{r}
study_t <- PL.clean %>% 
  group_by(month) %>% 
  summarize(mean_temp = mean(temp) %>% round(2)) 
head(study_t)

study_p <- PL.clean %>%
  group_by(month) %>%
  summarize(total_precip= mean(precip), sd_precip = sd(precip), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip - se_precip, upper = total_precip + se_precip %>% round(2)) 
head(study_p)
```




```{r}
p.climograph <- ggplot() +
  geom_boxplot(data = PL.clean, aes(x = month, y = temp), fill = "red3", alpha = 0.7) +
  geom_path(data = study_p, aes(x = month, y = total_precip-75, group=1), col= "steelblue4", size = 1) +
  geom_ribbon(data = study_p, aes(x = month, ymin = (lower-75), ymax = (upper-75), group=1), fill = "steelblue4", alpha = 0.2) +
  scale_y_continuous("Temperature (°C)", sec.axis = sec_axis(~ . + 75, name = "Precipitation (mm)")) +
  scale_x_discrete("Month") +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

p.climograph

#ggsave('p.climograph.pdf', p.climograph, units = 'cm', width = 20, height = 12)
```



```{r}
ggplot(data = PL.temp4, aes(x=year, y = temp)) +
  geom_line(alpha = 0.6) +
  facet_grid(month~., scales = "free") + 
  ylab("Mean Monthly Temperature (°C)") +
  theme_classic() 

ggplot(data = PL.precip4, aes(x=year, y = precip)) +
  geom_line(alpha = 0.6) +
  facet_grid(month~., scales = "free") + 
  ylab("Total Monthly Precipitation (mm)") +
  theme_classic() 

```

```{r}
ggplot(data = PL.temp4, aes(x=year, y = temp)) +
  geom_point(aes(col=source), size=0.3) +
  geom_line(alpha = 0.6) +
  facet_grid(month~., scales = "free") + 
  ylab("Mean Monthly Temperature (°C)") +
  theme_classic() 

ggplot(data = PL.temp4, aes(x=year, y = temp)) +
  geom_point(aes(col=source), size=0.3) +
  facet_grid(month~source, scales = "free") + 
  ylab("Mean Monthly Temperature (°C)") +
  theme_classic() 

ggplot(data = PL.precip4, aes(x=year, y = precip)) +
  geom_point(aes(col=source), size=0.3) +
  geom_line(alpha = 0.6) +
  facet_grid(month~., scales = "free") + 
  ylab("Total Monthly Precipitation (mm)") +
  theme_classic() 

ggplot(data = PL.precip4, aes(x=year, y = precip)) +
  geom_point(aes(col=source), size=0.3) +
  facet_grid(month~source, scales = "free") + 
  ylab("Total Monthly Precipitation (mm)") +
  theme_classic() 
```

```{r}
ggplot(data = PL.temp4, aes(x = month, y = temp)) + 
  geom_boxplot(alpha = 0.7, outlier.shape=NA) + 
  geom_jitter(aes(col=source), size=0.6, alpha=0.5) +
  ylab("Mean Monthly Temperature (°C)") +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

ggplot(data = PL.precip4, aes(x = month, y = precip)) + 
  geom_boxplot(alpha = 0.7, outlier.shape=NA) + 
  geom_jitter(aes(col=source), size=0.6, alpha=0.5) +
  ylab("Total Monthly Precipitation (mm)") +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```



# Annual Data

## Clean Record

```{r}
clean_temp_annual <- PL.clean %>% group_by(year) %>% summarise(mean_temp = mean(temp))
clean_precip_annual <- PL.clean %>% group_by(year) %>% summarise(mean_precip = mean(precip))
```

```{r}
ggplot() +
  geom_line(data=clean_temp_annual, aes(x = year, y = mean_temp), colour="lightsalmon4") +
  #Aesthetics
  xlab("Year") +
  ylab("Mean Annual Temperature (°C)") +
  scale_x_continuous(limits = c(1890, 2020), breaks=seq(1900,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 


ggplot() +
  geom_line(data=clean_precip_annual, aes(x = year, y = mean_precip), colour="lightsalmon4") +
  xlab("Year") +
  ylab("Total Annual Precipitation (mm)") +
  scale_x_continuous(limits = c(1890, 2020), breaks=seq(1900,2020,20)) +
 theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```


## Individual Records

```{r}
london_temp_annual <- london_temp %>% filter(month == "Annual")
london_precip_annual <- london_precip %>% filter(month == "Annual")
pinery_temp_annual <- pinery_temp_mean %>% group_by(year) %>% summarise(mean_temp = mean(mean_temp_year))
pinery_precip_annual <- pinery_precip_mean %>% group_by(year) %>% summarise(total_precip = sum(total_precip_year))
```

```{r}
london_temp_2 <- london_temp_annual %>% select(-month)
london_temp_2$location <- "London"
pinery_temp_annual$location <- "Pinery"
annual_temp <- rbind(london_temp_2,pinery_temp_annual) #join data sets

london_precip_2 <- london_precip_annual %>% select(-month)
london_precip_2$location <- "London"
pinery_precip_annual$location <- "Pinery"
annual_precip <- rbind(london_precip_2,pinery_precip_annual) #join data sets
```

```{r}
annual_temp_plot <- ggplot() +
  # London Data
  geom_line(data=london_temp_annual, aes(x = year, y = mean_temp), colour = "lightseagreen", linetype = "dashed", alpha = 0.8) +
  # Pinery Data
  geom_line(data=pinery_temp_annual, aes(x = year, y = mean_temp), colour="lightsalmon4") +
  #Aesthetics
  xlab("Year") +
  ylab("Mean Annual Temperature (°C)") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  #ggtitle("Average Annual Temperature in \n Southwestern Ontario from 1883 to 2017") + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

annual_temp_plot

#ggsave('annual_temp.pdf', annual_temp_plot, units = 'cm', width = 15, height = 10)

annual_precip_plot <- ggplot() +
  # London Data
  geom_line(data=london_precip_annual, aes(x = year, y = total_precip), colour = "lightseagreen", linetype = "dashed", alpha = 0.8) +
   # geom_smooth(data=london_precip_annual, aes(x = year, y = total_precip), se = TRUE, 
             # method = 'lm', size = 0.8, colour = "black", fill = "blue") +
  # Pinery Data
  geom_line(data=pinery_precip_annual, aes(x = year, y = total_precip, group = 1),  colour="lightsalmon4")+
   # geom_smooth(data=pinery_precip_annual, aes(x = year, y = total_precip), se = TRUE, 
            #  method = 'lm', size = 0.8, colour = "black", fill = "darkgreen") +
  #Aesthetics
  xlab("Year") +
  ylab("Total Annual Precipitation (mm)") +
  scale_x_continuous(breaks=seq(1880,2020,20)) +
  #ggtitle("Annual Precipitation in Southwestern Ontario \nfrom 1883 to 2017") + 
  scale_colour_manual(name="Location", breaks = c("London", "Pinery"),
                      values = c("blue","darkgreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

annual_precip_plot

#ggsave('annual_precip.pdf', annual_precip_plot, units = 'cm', width = 15, height = 10)

annual_plot <- grid.arrange(annual_temp_plot, annual_precip_plot, ncol= 2)
#ggsave('annual_plot.pdf', annual_plot, units = 'cm', width = 23, height = 8)
```


# Seasonal Data

```{r}
winter_t <- pinery_temp %>% filter(month == 1 | month == 2 | month == 3)
winter_t$season <- "winter"
spring_t <- pinery_temp %>% filter(month == 4 | month == 5 | month == 6)
spring_t$season <- "spring"
summer_t <- pinery_temp %>% filter(month == 7 | month == 8 | month == 9)
summer_t$season <- "summer"
fall_t <- pinery_temp %>% filter(month == 10 | month == 11 | month == 12)
fall_t$season <- "fall"
pinery_temp <- rbind(winter_t, spring_t, summer_t, fall_t)
pinery_temp$season <- as.factor(pinery_temp$season) #Make a factor 
str(pinery_temp)

pinery_temp_station <- pinery_temp %>% 
 group_by(year, season, station) %>% 
 summarise(mean_temp = mean(mean_temp))

pinery_temp_season <- pinery_temp_station %>%
  group_by(year, season) %>%
  summarise(mean_temp_season = mean(mean_temp) %>% round(2), sd_temp = sd(mean_temp) %>% round(2), n = n()) %>% 
  mutate(se_temp = (sd_temp / sqrt(n)) %>% round(2), lower = mean_temp_season - se_temp, upper = mean_temp_season + se_temp) 


winter_p <- pinery_precip %>% filter(month == 1 | month == 2 | month == 3)
winter_p$season <- "winter"
spring_p <- pinery_precip %>% filter(month == 4 | month == 5 | month == 6)
spring_p$season <- "spring"
summer_p <- pinery_precip %>% filter(month == 7 | month == 8 | month == 9)
summer_p$season <- "summer"
fall_p <- pinery_precip %>% filter(month == 10 | month == 11 | month == 12)
fall_p$season <- "fall"
pinery_precip <- rbind(winter_p, spring_p, summer_p, fall_p)
pinery_precip$season <- as.factor(pinery_precip$season) #Make a factor 

pinery_precip_station <- pinery_precip %>%
  group_by(year, season, station) %>%
  summarise(total_precip = mean(total_precip))

pinery_precip_season <- pinery_precip_station %>% 
  group_by(year, season) %>% 
  summarise(total_precip_season = mean(total_precip) %>% round(2), sd_precip = sd(total_precip) %>% round(2), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip_season - sd_precip, upper = total_precip_season + sd_precip %>% round(2)) 

```

```{r}
winter_tl <- london_temp %>% filter(month == "Jan" | month == "Feb" | month == "Mar")
winter_tl$season <- "winter"
spring_tl <- london_temp %>% filter(month == "Apr" | month == "May" | month == "Jun")
spring_tl$season <- "spring"
summer_tl <- london_temp %>% filter(month == "Jul" | month == "Aug" | month == "Sep")
summer_tl$season <- "summer"
fall_tl <- london_temp %>% filter(month == "Oct" | month == "Nov" | month == "Dec")
fall_tl$season <- "fall"
london_temp <- rbind(winter_tl, spring_tl, summer_tl, fall_tl)
london_temp$season <- as.factor(london_temp$season) #Make a factor 

london_temp_season <- london_temp %>% 
  group_by(year, season) %>% 
  summarise(mean_temp_season = mean(mean_temp))


winter_pl <- london_precip %>% filter(month == "Jan" | month == "Feb" | month == "Mar")
winter_pl$season <- "winter"
spring_pl <- london_precip %>% filter(month == "Apr" | month == "May" | month == "Jun")
spring_pl$season <- "spring"
summer_pl <- london_precip %>% filter(month == "Jul" | month == "Aug" | month == "Sep")
summer_pl$season <- "summer"
fall_pl <- london_precip %>% filter(month == "Oct" | month == "Nov" | month == "Dec")
fall_pl$season <- "fall"
london_precip <- rbind(winter_pl, spring_pl, summer_pl, fall_pl)
london_precip$season <- as.factor(london_precip$season) #Make a factor 

london_precip_season <- london_precip %>% 
  group_by(year, season) %>% 
  summarise(total_precip_season = mean(total_precip))
```
```{r}
london_temp_season$season <- factor(london_temp_season$season, levels = c("winter", "spring", "summer", "fall"))
pinery_temp_season$season <- factor(pinery_temp_season$season, levels = c("winter", "spring", "summer", "fall"))
```


```{r}
pinery_temp_season$location <- "Pinery"
london_temp_season$location <- "London"
season_temp <- rbind(pinery_temp_season, london_temp_season)

pinery_precip_season$location <- "Pinery"
london_precip_season$location <- "London"
season_precip <- rbind(pinery_precip_season, london_precip_season)

season_temp$location <- factor(season_temp$location, levels = c("Pinery", "London"))
season_precip$location <- factor(season_precip$location, levels = c("Pinery", "London"))
```



## Temperature

```{r}
w_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "winter")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "spring")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "summer")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "fall")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_temp_plot <- grid.arrange(w_temp, sp_temp, su_temp, fa_temp, nrow = 2, ncol = 2)
#ggsave('seasonal_temperature.pdf', seasonal_temp_plot, units = 'cm', width = 20, height = 12)
```



## Precipitation
```{r}
w_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "winter")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "spring")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "summer")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "fall")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_precip_plot <- grid.arrange(w_precip, sp_precip, su_precip, fa_precip, nrow = 2, ncol = 2)
#ggsave('seasonal_precipitation.pdf', seasonal_precip_plot, units = 'cm', width = 20, height = 12)
```



## Temperature Anomaly

https://tamino.wordpress.com/2018/07/29/why-use-temperature-anomaly/

Average defined as 1951-1980 (NASA standard)

```{r}
pinery_average <- season_temp %>% filter(location == "Pinery") %>% filter(year > 1950 & year < 1981)
pinery_average <- pinery_average %>% group_by(season) %>% summarise(pinery_average = mean(mean_temp_season))

london_average <- season_temp %>% filter(location == "London") %>% filter(year > 1950 & year < 1981)
london_average <- london_average %>% group_by(season) %>% summarise(london_average = mean(mean_temp_season))

pinery_anomaly <- right_join(season_temp, pinery_average)
pinery_anomaly <- mutate(pinery_anomaly, anomaly = mean_temp_season - pinery_average) %>% filter(location == "Pinery")

london_anomaly <- right_join(season_temp, london_average)
london_anomaly <- mutate(london_anomaly, anomaly = mean_temp_season - london_average) %>% filter(location == "London")

season_anomaly <- full_join(pinery_anomaly, london_anomaly) %>% select(-pinery_average, -london_average)
```

```{r}
w_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "winter")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "spring")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "summer")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "fall")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_anomaly_plot <- grid.arrange(w_anomaly, sp_anomaly, su_anomaly, fa_anomaly, nrow = 2, ncol = 2)
#ggsave('seasonal_anomaly.pdf', seasonal_anomaly_plot, units = 'cm', width = 20, height = 12)
```



# Save Outputs

```{r}
write.csv(PL.clean, "outputs/full_climate.csv")
write.csv(PL.temp4, "outputs/full_temp.csv")
write.csv(PL.precip4, "outputs/full_precip.csv")
write.csv(pinery_temp_mean, "outputs/pinery_monthly_temperature.csv")
write.csv(pinery_precip_mean, "outputs/pinery_monthly_precipitation.csv")
write.csv(season_temp, file = "outputs/pinery_season_temperature.csv")
write.csv(season_precip, file = "outputs/pinery_season_precipitation.csv")
write.csv(pinery_temp_annual, file = "outputs/pinery_annual_temp.csv")
```


# Reproducibility
```{r}
Sys.time()
git2r::repository()
sessionInfo()
```






