---
title: "Instrumental Climate Data"
author: "Jen Baron, j.baron@alumni.ubc.ca, UBC Tree Ring Lab"
date: 'August 8, 2019'
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
---

To Do:

- Fill missing data
- Update Pinery Climate (96 - present; Thedford & our data)
- Growing season T (May - Sept)
- Vapour pressure defecit (growing season T & previous Aug-Oct, Williams et al.)


# Packages
```{r}
library(ggplot2) #for plotting
library(tidyr)
library(dplyr)
library(GGally)
library(gridExtra)
```


# Read in Data

The Pinery dataset was compiled in excel from monthly mean temperature and total precipitation data using open access Environment Canada records for stations within 30km of the Pinery with at least 10 consecutive years of data (http://climate.weather.gc.ca/historical_data/search_historic_data_e.html). 

The London datasets are homogenized data from Environment Canada that were converted to tidy format in excel. 

```{r}
pinery <- read.csv("data/pinery_climate.csv")
pinery$station <- as.factor(pinery$station) #Make a factor 
str(pinery)
london_precip <- read.csv("data/london_precip.csv")
london_temp <- read.csv("data/london_temp.csv")
london_precip$month <- as.factor(london_precip$month) #Make a factor 
london_temp$month <- as.factor(london_temp$month) #Make a factor 
```


# Create Regional Data

## Clean Data

Here I go through each station for the temp and precip data, removing years where all 12 months aren't present so that I can later calculate the mean values for each year. I did this manually a while ago and there is probably a more efficient way to do it, however it works and it's already done so I left it as is. 
```{r}
#Arkona
Arkona <- pinery %>% filter(station == "Arkona") #Filter out station of interest

Arkona_T <- Arkona %>% select(-total_precip) %>% filter(!is.na(Arkona$mean_temp)==T) #remove precipitation data and "NA"
table(Arkona_T$year) #Generate table after cleaning
Arkona_T2 <- Arkona_T %>% filter(year != 1882 & year != 1883 & 
                                   year != 1884 & year != 1893 & year != 1900 & year != 1902 & year != 1915) #remove years with < 12 months
table(Arkona_T2$year) #Generate table after cleaning

Arkona_P <- Arkona %>% select(-mean_temp) %>% filter(!is.na(Arkona$total_precip)==T) #remove temperature data and "NA"
#table(Arkona_P$year) #show how many observations in each year
Arkona_P2 <- Arkona_P %>% filter(year != 1882 & year != 1893 & year != 1902 & year != 1915)
#table(Arkona_P2$year) #Generate table after cleaning

#Centralia A
CentraliaA <- pinery %>% filter(station == "Centralia A")

CentraliaA_T <- CentraliaA %>% select(-total_precip) %>% filter(!is.na(CentraliaA$mean_temp)==T) #remove "NA"
#table(CentraliaA_T$year)  
CentraliaA_T2 <- CentraliaA_T %>% filter(year != 1945 & year != 1946 & year != 1947 & year != 1966)
#table(CentraliaA_T2$year)  

CentraliaA_P <- CentraliaA %>% select(-mean_temp) %>% filter(!is.na(CentraliaA$total_precip)==T) #remove "NA"
#table(CentraliaA_P$year)  
CentraliaA_P2 <- CentraliaA_P %>% filter(year != 1946 & year != 1947 & year != 1966)
#table(CentraliaA_P2$year)  

#Exeter
Exeter <- pinery %>% filter(station == "Exeter")

Exeter_T <- Exeter %>% select(-total_precip) %>% filter(!is.na(Exeter$mean_temp)==T) #remove "NA"
#table(Exeter_T$year)  
Exeter_T2 <- Exeter_T %>% filter(year != 1967 & year != 1976 & year != 1985)
#table(Exeter_T2$year)  

Exeter_P <- Exeter %>% select(-mean_temp) %>% filter(!is.na(Exeter$total_precip)==T) #remove "NA"
#table(Exeter_P$year)  
Exeter_P2 <- Exeter_P %>% filter(year != 1961 & year != 1962 & year != 1964 & year != 1976 & year != 1985)
#table(Exeter_P2$year) 

#Forest
Forest <- pinery %>% filter(station == "Forest")

Forest_T <- Forest %>% select(-total_precip) %>% filter(!is.na(Forest$mean_temp)==T) #remove "NA"
#table(Forest_T$year)  
Forest_T2 <- Forest_T %>% filter(year != 1924 & year != 1940 & year != 1962 & year != 1963)
#table(Forest_T2$year) 

Forest_P <- Forest %>% select(-mean_temp) %>% filter(!is.na(Forest$total_precip)==T) #remove "NA"
#table(Forest_P$year)  
Forest_P2 <- Forest_P %>% filter(year > 1933 & year != 1937 & year!= 1943 & year != 1955 & year != 1960 & year < 1960)
#table(Forest_P2$year) 

#Lucan
Lucan <- pinery %>% filter(station == "Lucan")

Lucan_T <- Lucan %>% select(-total_precip) %>% filter(!is.na(Lucan$mean_temp)==T) #remove "NA"
#table(Lucan_T$year)  
Lucan_T2 <- Lucan_T %>% filter(year != 1915 & year != 1924 & year != 1948 & year != 1949 & year != 1950
                               & year != 1951 & year != 1952 & year != 1953 & year != 1954 & year != 1955
                               & year != 1956 & year != 1957 & year != 1958 & year != 1959 & year != 1961)
#table(Lucan_T2$year) 

Lucan_P <- Lucan %>% select(-mean_temp) %>% filter(!is.na(Lucan$total_precip)==T) #remove "NA"
#table(Lucan_P$year)  
Lucan_P2 <- Lucan_P %>% filter(year != 1915 & year != 1923 & year != 1943 & year < 1948)
#table(Lucan_P2$year)

#Pinery
Pinery <- pinery %>% filter(station == "Pinery")

Pinery_T <- Pinery %>% select(-total_precip) %>% filter(!is.na(Pinery$mean_temp)==T) #remove "NA"
#table(Pinery_T$year)  
Pinery_T2 <- Pinery_T %>% filter(year != 1979 & year != 1984)
#table(Pinery_T2$year) 

Pinery_P <- Pinery %>% select(-mean_temp) %>% filter(!is.na(Pinery$total_precip)==T) #remove "NA"
#table(Pinery_P$year)  
Pinery_P2 <- Pinery_P %>% filter(year != 1979 & year != 1984)
#table(Pinery_P2$year)

#Strathrow-Mullifary
Strathroy <- pinery %>% filter(station == "Strathroy-Mullifary")

Strathroy_T <- Strathroy %>% select(-total_precip) %>% filter(!is.na(Strathroy$mean_temp)==T) #remove "NA"
#table(Strathroy_T$year)  
Strathroy_T2 <- Strathroy_T %>% filter(year != 2009 & year != 2010 & year != 2013 & year != 2014 & year != 2015 & year < 2017)
#table(Strathroy_T2$year) 

Strathroy_P <- Strathroy %>% select(-mean_temp) %>% filter(!is.na(Strathroy$total_precip)==T) #remove "NA"
#table(Strathroy_P$year)  
Strathroy_P2 <- Strathroy_P %>% filter(year != 2009 & year != 2010 & year != 2012 & year != 2013 
                                       & year != 2014 & year != 2015 & year < 2017)
#table(Strathroy_P2$year)

#Thedford (no temp)
Thedford <- pinery %>% filter(station == "Thedford")

Thedford_P <- Thedford %>% select(-mean_temp) %>% filter(!is.na(Thedford$total_precip)==T) #remove "NA"
#table(Thedford_P$year)  
Thedford_P2 <- Thedford_P %>% filter(year != 1883 & year != 1897)
#table(Thedford_P2$year)

#Watford (can't use precipitation data)
Watford <- pinery %>% filter(station == "Watford")

Watford_T <- Watford %>% select(-total_precip) %>% filter(!is.na(Watford$mean_temp)==T) #remove "NA"
#table(Watford_T$year)  
Watford_T2 <- Watford_T %>% filter(year == 1927 | year == 1928)
#table(Watford_T2$year) 

Watford_P <- Watford %>% select(-mean_temp) %>% filter(!is.na(Watford$total_precip)==T) #remove "NA"
#table(Watford_P$year)  
```


EE Re-combine Data
```{r}
pinery_temp <- rbind(Arkona_T2, CentraliaA_T2, Exeter_T2, Forest_T2, Lucan_T2, Pinery_T2, Strathroy_T2, Watford_T2)
pinery_precip <- rbind(Arkona_P2, CentraliaA_P2, Exeter_P2, Forest_P2, Lucan_P2, Pinery_P2, Strathroy_P2, Thedford_P2)
```


Ensure the cleaning worked by plotting histograms of the number of observations (months) in each year (should only be multiples of 12)
```{r}
#histogram (temperature)
graph <- ggplot(pinery_temp, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
  geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
    geom_hline(yintercept = 24, linetype="solid", color = "red") + 
    geom_hline(yintercept = 36, linetype="solid", color = "red") + 
  theme_bw()
graph

#histogram (precipitation)
graph <- ggplot(pinery_precip, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey") +
   geom_hline(yintercept = 12, linetype="solid", color = "red") + #shows 12 months
    geom_hline(yintercept = 24, linetype="solid", color = "red") + 
  theme_bw()
graph
```

## Calculate Average (month-year)

Calculate one mean temperature and total precipitation for each year/month where more than one station of data is available

```{r}
#Temperature
str(pinery_temp)

pinery_temp_mean <- pinery_temp %>%
  group_by(year,month) %>%
  summarise(mean_temp_year = mean(mean_temp), sd_temp = sd(mean_temp), n = n()) %>%
  mutate(se_temp = sd_temp / sqrt(n), lower = (mean_temp_year - se_temp), upper = (mean_temp_year + se_temp)) %>% round(2)
pinery_temp_mean$month <- as.factor(pinery_temp_mean$month)

#Precipitation
pinery_precip_mean <- pinery_precip %>%
  group_by(year,month) %>%
  summarise(total_precip_year = mean(total_precip), sd_precip = sd(total_precip), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip_year - sd_precip, upper = total_precip_year + sd_precip) %>% round(2)

head(pinery_precip_mean)
```

```{r}
ggplot(pinery_precip_mean, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey")

ggplot(pinery_temp_mean, aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey")
```



## Fill Empty Years

```{r}
year <- c(1885:2017) #create time vector
month <- c(1:12)
str(time)
```

Fill empty months-years with data from London

```{r}
p.t <- pinery_temp_mean %>% select(year, month, mean_temp_year)

l.t <- london_temp %>% filter(month == "Jan" |
                                month == "Feb" |
                                month == "Mar" |
                                month == "Apr" |
                                month == "May" |
                                month == "Jun" |
                                month == "Jul" |
                                month == "Aug" |
                                month == "Sep" |
                                month == "Oct" |
                                month == "Nov" |
                                month == "Dec") %>% 
  mutate(month = recode(month, "Jan"="1", 
                                     "Feb"="2", 
                                     "Mar"="3",
                                     "Apr"="4",
                                     "May"="5", 
                                     "Jun"="6", 
                                     "Jul"="7",
                                     "Aug"="8",
                                     "Sep"="9",
                                     "Oct"="10",
                                     "Nov"="11",
                                     "Dec"="12"))

```


```{r}
PL.temp <- full_join(l.t, pinery_temp_month)
```

```{r}
ggplot((l.t %>% na.omit), aes(x = year)) +
  geom_histogram(position = 'identity', binwidth = 1, color = "black", fill = "grey")
```




Select all rows with regional data - source = Pinery

Select all rows with NA - source = London

```{r}
PL.1 <- PL.temp %>% 
  filter(is.na(mean_temp_year))
PL.1$source <- "London"
PL.1$temp <- PL.1$mean_temp

PL.2 <- PL.temp %>%
  na.omit(cols=c("mean_temp_year"))
PL.2$source <- "Pinery"
PL.2$temp <- PL.2$mean_temp_year

PL.temp2 <- rbind(PL.1, PL.2) 
```

Also missing 1940:7
Create data frame with all months & year to ensure none are missing*

Fill leftover NAs with climateNA values?

```{r}
PL.temp2 %>% filter(is.na(temp))
```








# Monthly Data

## Climograph

```{r}
study_t <- pinery_temp_mean %>% 
  group_by(month) %>% 
  summarize(mean_temp = mean(mean_temp_year)) %>% round(2)
study_t

study_p <- pinery_precip_mean %>%
  group_by(month) %>%
  summarize(total_precip= mean(total_precip_year), sd_precip = sd(total_precip_year), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip - se_precip, upper = total_precip + se_precip) %>% round(2)
study_p
```

```{r}
ggplot() + 
geom_col(data = study_t, aes(x = month, y = mean_temp), fill = "red3", alpha = 0.7) + 
geom_line(data = study_p, aes(x = month, y = total_precip-80), col= "steelblue4", size = 1.5) +
  scale_y_continuous("Temperature (°C)", sec.axis = sec_axis(~ . + 80, name = "Precipitation (mm)")) +
    geom_ribbon(data = study_p, aes(x = month, ymin = (lower-80), ymax = (upper-80)), alpha = 0.3) +
  scale_x_continuous("Month", breaks = c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
```



```{r}
p.climograph <- ggplot() +
  geom_boxplot(data = pinery_temp_mean, aes(x = month, y = mean_temp_year, group = month), fill = "red3", alpha = 0.7) +
  geom_line(data = study_p, aes(x = month, y = total_precip-75), col= "steelblue4", size = 1) +
  geom_ribbon(data = study_p, aes(x = month, ymin = (lower-75), ymax = (upper-75)), fill = "steelblue4", alpha = 0.2) +
  scale_y_continuous("Temperature (°C)", sec.axis = sec_axis(~ . + 75, name = "Precipitation (mm)")) +
  scale_x_continuous("Month", breaks = c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

p.climograph

#ggsave('p.climograph.pdf', p.climograph, units = 'cm', width = 20, height = 12)
```


# Annual Data

```{r}
london_temp_annual <- london_temp %>% filter(month == "Annual")
london_precip_annual <- london_precip %>% filter(month == "Annual")
pinery_temp_annual <- pinery_temp_mean %>% group_by(year) %>% summarise(mean_temp = mean(mean_temp_year))
pinery_precip_annual <- pinery_precip_mean %>% group_by(year) %>% summarise(total_precip = sum(total_precip_year))
```

```{r}
london_temp_2 <- london_temp_annual %>% select(-month)
london_temp_2$location <- "London"
pinery_temp_annual$location <- "Pinery"
annual_temp <- rbind(london_temp_2,pinery_temp_annual) #join data sets

london_precip_2 <- london_precip_annual %>% select(-month)
london_precip_2$location <- "London"
pinery_precip_annual$location <- "Pinery"
annual_precip <- rbind(london_precip_2,pinery_precip_annual) #join data sets
```


Plot Temperature
```{r}
annual_temp_plot <- ggplot() +
  # London Data
  geom_line(data=london_temp_annual, aes(x = year, y = mean_temp), colour = "lightseagreen", linetype = "dashed", alpha = 0.8) +
  # Pinery Data
  geom_line(data=pinery_temp_annual, aes(x = year, y = mean_temp), colour="lightsalmon4") +
  #Aesthetics
  xlab("Year") +
  ylab("Mean Annual Temperature (°C)") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  #ggtitle("Average Annual Temperature in \n Southwestern Ontario from 1883 to 2017") + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

annual_temp_plot

#ggsave('annual_temp.pdf', annual_temp_plot, units = 'cm', width = 15, height = 10)

```

Plot recent temperature
```{r}
pinery_temp_annual_recent <- pinery_temp_annual %>% filter(year > 1949)

annual_temp_plot2 <- ggplot() +
  # Pinery Data
  geom_line(data=pinery_temp_annual_recent, aes(x = year, y = mean_temp)) +
  geom_point(data=pinery_temp_annual_recent, aes(x = year, y = mean_temp), size = 1)  +
  #Aesthetics
  xlab("Year") +
  ylab("Temperature (°C)") +
  scale_x_continuous(breaks=seq(1950,2020,10)) +
  #ggtitle("Average Annual Temperature in \n Southwestern Ontario from 1883 to 2017") + 
  scale_colour_manual(name="Location", breaks = c("London", "Pinery"),
                      values = c("blue","darkgreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
#ggsave('annual_temp.pdf', annual_temp_plot, units = 'cm', width = 20, height = 12)
annual_temp_plot2
```

Plot Precipitation
```{r}
annual_precip_plot <- ggplot() +
  # London Data
  geom_line(data=london_precip_annual, aes(x = year, y = total_precip), colour = "lightseagreen", linetype = "dashed", alpha = 0.8) +
   # geom_smooth(data=london_precip_annual, aes(x = year, y = total_precip), se = TRUE, 
             # method = 'lm', size = 0.8, colour = "black", fill = "blue") +
  # Pinery Data
  geom_line(data=pinery_precip_annual, aes(x = year, y = total_precip, group = 1),  colour="lightsalmon4")+
   # geom_smooth(data=pinery_precip_annual, aes(x = year, y = total_precip), se = TRUE, 
            #  method = 'lm', size = 0.8, colour = "black", fill = "darkgreen") +
  #Aesthetics
  xlab("Year") +
  ylab("Total Annual Precipitation (mm)") +
  scale_x_continuous(breaks=seq(1880,2020,20)) +
  #ggtitle("Annual Precipitation in Southwestern Ontario \nfrom 1883 to 2017") + 
  scale_colour_manual(name="Location", breaks = c("London", "Pinery"),
                      values = c("blue","darkgreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

annual_precip_plot

#ggsave('annual_precip.pdf', annual_precip_plot, units = 'cm', width = 15, height = 10)
```

```{r}
annual_plot <- grid.arrange(annual_temp_plot, annual_precip_plot, ncol= 2)
#ggsave('annual_plot.pdf', annual_plot, units = 'cm', width = 23, height = 8)
```


# Seasonal Data

```{r}
winter_t <- pinery_temp %>% filter(month == 1 | month == 2 | month == 3)
winter_t$season <- "winter"
spring_t <- pinery_temp %>% filter(month == 4 | month == 5 | month == 6)
spring_t$season <- "spring"
summer_t <- pinery_temp %>% filter(month == 7 | month == 8 | month == 9)
summer_t$season <- "summer"
fall_t <- pinery_temp %>% filter(month == 10 | month == 11 | month == 12)
fall_t$season <- "fall"
pinery_temp <- rbind(winter_t, spring_t, summer_t, fall_t)
pinery_temp$season <- as.factor(pinery_temp$season) #Make a factor 
str(pinery_temp)

pinery_temp_station <- pinery_temp %>% 
 group_by(year, season, station) %>% 
 summarise(mean_temp = mean(mean_temp))

pinery_temp_season <- pinery_temp_station %>%
  group_by(year, season) %>%
  summarise(mean_temp_season = mean(mean_temp) %>% round(2), sd_temp = sd(mean_temp) %>% round(2), n = n()) %>% 
  mutate(se_temp = (sd_temp / sqrt(n)) %>% round(2), lower = mean_temp_season - se_temp, upper = mean_temp_season + se_temp) 

head(pinery_temp_season)
```

```{r}
winter_p <- pinery_precip %>% filter(month == 1 | month == 2 | month == 3)
winter_p$season <- "winter"
spring_p <- pinery_precip %>% filter(month == 4 | month == 5 | month == 6)
spring_p$season <- "spring"
summer_p <- pinery_precip %>% filter(month == 7 | month == 8 | month == 9)
summer_p$season <- "summer"
fall_p <- pinery_precip %>% filter(month == 10 | month == 11 | month == 12)
fall_p$season <- "fall"
pinery_precip <- rbind(winter_p, spring_p, summer_p, fall_p)
pinery_precip$season <- as.factor(pinery_precip$season) #Make a factor 

pinery_precip_station <- pinery_precip %>%
  group_by(year, season, station) %>%
  summarise(total_precip = mean(total_precip))

pinery_precip_season <- pinery_precip_station %>% 
  group_by(year, season) %>% 
  summarise(total_precip_season = mean(total_precip) %>% round(2), sd_precip = sd(total_precip) %>% round(2), n = n()) %>%
  mutate(se_precip = sd_precip / sqrt(n), lower = total_precip_season - sd_precip, upper = total_precip_season + sd_precip %>% round(2)) 

head(pinery_precip_season)
```

```{r}
winter_tl <- london_temp %>% filter(month == "Jan" | month == "Feb" | month == "Mar")
winter_tl$season <- "winter"
spring_tl <- london_temp %>% filter(month == "Apr" | month == "May" | month == "Jun")
spring_tl$season <- "spring"
summer_tl <- london_temp %>% filter(month == "Jul" | month == "Aug" | month == "Sep")
summer_tl$season <- "summer"
fall_tl <- london_temp %>% filter(month == "Oct" | month == "Nov" | month == "Dec")
fall_tl$season <- "fall"
london_temp <- rbind(winter_tl, spring_tl, summer_tl, fall_tl)
london_temp$season <- as.factor(london_temp$season) #Make a factor 
str(london_temp)

london_temp_season <- london_temp %>% 
  group_by(year, season) %>% 
  summarise(mean_temp_season = mean(mean_temp))
```

```{r}
winter_pl <- london_precip %>% filter(month == "Jan" | month == "Feb" | month == "Mar")
winter_pl$season <- "winter"
spring_pl <- london_precip %>% filter(month == "Apr" | month == "May" | month == "Jun")
spring_pl$season <- "spring"
summer_pl <- london_precip %>% filter(month == "Jul" | month == "Aug" | month == "Sep")
summer_pl$season <- "summer"
fall_pl <- london_precip %>% filter(month == "Oct" | month == "Nov" | month == "Dec")
fall_pl$season <- "fall"
london_precip <- rbind(winter_pl, spring_pl, summer_pl, fall_pl)
london_precip$season <- as.factor(london_precip$season) #Make a factor 
str(london_precip)

london_precip_season <- london_precip %>% 
  group_by(year, season) %>% 
  summarise(total_precip_season = mean(total_precip))
```

Make a factor

```{r}
london_temp_season$season <- factor(london_temp_season$season, levels = c("winter", "spring", "summer", "fall"))
pinery_temp_season$season <- factor(pinery_temp_season$season, levels = c("winter", "spring", "summer", "fall"))
```

## Combine London & Pinery Records

```{r}
pinery_temp_season$location <- "Pinery"
london_temp_season$location <- "London"
season_temp <- rbind(pinery_temp_season, london_temp_season)

pinery_precip_season$location <- "Pinery"
london_precip_season$location <- "London"
season_precip <- rbind(pinery_precip_season, london_precip_season)

season_temp$location <- factor(season_temp$location, levels = c("Pinery", "London"))
season_precip$location <- factor(season_precip$location, levels = c("Pinery", "London"))
```


Plot Temperature
```{r}
w_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "winter")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "spring")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "summer")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_temp <- ggplot() +
  geom_line(data=(season_temp %>% filter(season == "fall")), aes(x = year, y = mean_temp_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature (°C)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
    scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_temp_plot <- grid.arrange(w_temp, sp_temp, su_temp, fa_temp, nrow = 2, ncol = 2)
#ggsave('seasonal_temperature.pdf', seasonal_temp_plot, units = 'cm', width = 20, height = 12)
```





Plot Precipitation
```{r}
w_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "winter")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "spring")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "summer")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_precip <- ggplot() +
  geom_line(data=(season_precip %>% filter(season == "fall")), aes(x = year, y = total_precip_season, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Precipitation (mm)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
      scale_color_manual(values = c("lightsalmon4", "lightseagreen")) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_precip_plot <- grid.arrange(w_precip, sp_precip, su_precip, fa_precip, nrow = 2, ncol = 2)
#ggsave('seasonal_precipitation.pdf', seasonal_precip_plot, units = 'cm', width = 20, height = 12)
```

Pink = London
Blue = Pinery


## Temperature Anomaly

https://tamino.wordpress.com/2018/07/29/why-use-temperature-anomaly/

Average defined as 1951-1980 (NASA standard)

```{r}
pinery_average <- season_temp %>% filter(location == "Pinery") %>% filter(year > 1950 & year < 1981)
pinery_average <- pinery_average %>% group_by(season) %>% summarise(pinery_average = mean(mean_temp_season))

london_average <- season_temp %>% filter(location == "London") %>% filter(year > 1950 & year < 1981)
london_average <- london_average %>% group_by(season) %>% summarise(london_average = mean(mean_temp_season))
```


```{r}
pinery_anomaly <- right_join(season_temp, pinery_average)
pinery_anomaly <- mutate(pinery_anomaly, anomaly = mean_temp_season - pinery_average) %>% filter(location == "Pinery")

london_anomaly <- right_join(season_temp, london_average)
london_anomaly <- mutate(london_anomaly, anomaly = mean_temp_season - london_average) %>% filter(location == "London")

season_anomaly <- full_join(pinery_anomaly, london_anomaly) %>% select(-pinery_average, -london_average)
```
```{r}
w_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "winter")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "spring")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "summer")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "fall")), aes(x = year, y = anomaly, col = location, linetype = location)) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1880, 2020), breaks=seq(1880,2020,20)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_anomaly_plot <- grid.arrange(w_anomaly, sp_anomaly, su_anomaly, fa_anomaly, nrow = 2, ncol = 2)
#ggsave('seasonal_anomaly.pdf', seasonal_anomaly_plot, units = 'cm', width = 20, height = 12)
```

Plot recent temperature anomaly

```{r}
w_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "winter", year > 1949, location == "Pinery")), aes(x = year, y = anomaly)) +
  geom_point(data=(season_anomaly %>% filter(season == "winter", year > 1949, location == "Pinery")), aes(x = year, y = anomaly), size = 0.5) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Winter") +
  scale_x_continuous(limits = c(1950, 2020), breaks=seq(1950,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

sp_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "spring", year > 1949, location == "Pinery")), aes(x = year, y = anomaly)) +
 geom_point(data=(season_anomaly %>% filter(season == "spring", year > 1949, location == "Pinery")), aes(x = year, y = anomaly), size = 0.5) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Spring") +
  scale_x_continuous(limits = c(1950, 2020), breaks=seq(1950,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

su_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "summer", year > 1949, location == "Pinery")), aes(x = year, y = anomaly)) +
  geom_point(data=(season_anomaly %>% filter(season == "summer", year > 1949, location == "Pinery")), aes(x = year, y = anomaly), size = 0.5) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Summer") +
  scale_x_continuous(limits = c(1950, 2020), breaks=seq(1950,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") 

fa_anomaly <- ggplot() +
  geom_line(data=(season_anomaly %>% filter(season == "fall", year > 1949, location == "Pinery")), aes(x = year, y = anomaly)) +
    geom_point(data=(season_anomaly %>% filter(season == "fall", year > 1949, location == "Pinery")), aes(x = year, y = anomaly), size = 0.5) +
  #Aesthetics
  xlab("Year") + ylab("Temperature Anomaly (°C)") + ggtitle("Fall") +
  scale_x_continuous(limits = c(1950, 2020), breaks=seq(1950,2020,10)) +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position = "none") 

seasonal_anomaly_plot <- grid.arrange(w_anomaly, sp_anomaly, su_anomaly, fa_anomaly, nrow = 2, ncol = 2)
```



# Save Outputs

```{r}
write.csv(pinery_temp_mean, "outputs/monthly_temperature.csv")
write.csv(pinery_precip_mean, "outputs/monthly_precipitation.csv")
#write.csv(season_temp, file = "outputs/season_temperature.csv")
#write.csv(season_precip, file = "outputs/season_precipitation.csv")
#write.csv(pinery_temp_annual, file = "outputs/annual_temp.csv")
```


# Reproducibility
```{r}
Sys.time()
git2r::repository()
sessionInfo()
```






